### 2020-05-29 text/article_vedrovice.Rmd
text and SI - methods

---
title           : "Supplement: Vedrovice burial analysis"
author          : "Petr Pajdla"
    # name          : "Petr Pajdla"
    # affiliation   : "Department of Archaeology and Museology,
    #                  Masaryk University, Brno, Czech Republic"
    # email         : "petr.pajdla@gmail.com"
date            : "`r format(Sys.time(), '%d %B, %Y')`"
abstract        : "Fuck me."
output :
    pdf_document:
        latex: pdflatex
        fig_caption: true
        number_sections: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "european-journal-of-archaeology.csl"
---

@@

@@

## Randomness of Artefact Combinations

To determine the non-randomness of various artefact combinations we conduct a
permutation test on the presence/absence data matrix.
The method based on a theory of generalized Monte Carlo tests was first
introduced to archaeology by @manly1996.
Original (observed) co-occurence matrix is permutated $n$ times with constant
row and column totals, i. e. both the count of artifacts in individual graves
and sum of the given type of artifacts are fixed.

To examine which associations of variables are infrequent or even not present
and which tend to appear together we conduct a randomization test

@@

Exceptionality index ($EI$) is used in order to induce relative exceptionality
of individual burials based on their attributes.
This approach is independent on a prior selection of attributes of importance
or *prestige*.
The concept was introduced by @demjan2015.

It is defined by @demjan2015[367] as a Root Mean Square (*quadratic mean*) of
PCA factor scores:
$$ EI_{attr_g}=\sqrt{\frac{1}{n}\sum_{i=1}^{n}s_i^2} $$
where $s_i$ are factor scores of individual graves. The mean value is
normalized to range $[0-1]$:

@@

>>>>

---
title           : "Vedrovice burial analysis"
author          : "Petr Pajdla"
    # name          : "Petr Pajdla"
    # affiliation   : "Department of Archaeology and Museology,
    #                  Masaryk University, Brno, Czech Republic"
    # email         : "petr.pajdla@gmail.com"
date            : "`r format(Sys.time(), '%d %B, %Y')`"
abstract        : "Fuck me."
output:
    bookdown::pdf_document2:
    # bookdown::word_document2:
        latex: pdflatex
        fig_caption: true
        number_sections: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "european-journal-of-archaeology.csl"
---

@@

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
```

@@

## Randomness of artefact combinations

To determine non-randomness of various artefact combinations we conduct a
permutation test on the presence/absence data matrix.
The method based on generalized Monte Carlo tests was first
introduced to archaeology by @manly1996.
Original (observed) co-occurence matrix is permutated $n$ times with constant
row and column totals, i. e. both the count of artifacts in individual graves
and sum of the given type of artifacts across the burials are fixed.

To examine which associations of variables are infrequent or even not present
and which tend to appear together co-occurences of observed and simulated data
are summarized using two statistics suggested by @manly1996.
The $S$ statistic is a measure of overall deviation between expected and observed
number of artefact occurences while
the $v$ statistic is a measure of difference between co-occurences of
individual artefact with other artifacts
[@manly1996, 475, for details see SI].

@@

Exceptionality index ($EI$) is used in order to induce relative exceptionality
of individual burials based on their attributes.
This approach is independent on a prior selection of attributes of importance
or *prestige*.
It was defined by @demjan2015[367] based on a Root Mean Square (*quadratic mean*) of
PCA factor scores:
$$ EI_{attr_g}=\sqrt{\frac{1}{n}\sum_{i=1}^{n}s_i^2} $$
where $s_i$ are factor scores of individual graves. The mean value is
normalized to range $[0-1]$:

@@

## Randomness of artefact combinations

```{r vtable}
vstat <- read_csv(here::here("data/temp", "v_statistics.csv")) %>% 
    select(full, v, p) %>% 
    mutate(v = round(v, 2), p = round(p, 2)) %>% 
    rename("variable" = full, "v statistic" = v, "p value" = p)

knitr::kable(
  vstat,
  caption = 'Fuuuuu'
)
```
